---
title: "Cell types affected in Parkinson's post mortem transcriptomes"
output: html_notebook
---

# Prep midbrain specific cell type specificity dataset

For the cell type specificity matrix, we only want to include cell types which could exist within the substantia nigra. 

```{r, eval=FALSE}
if (!require("remotes")) install.packages("remotes")
if (!require("pacman")) install.packages("pacman")
if (!require("BiocManager")) install.packages("BiocManager")
pacman::p_load("rhdf5","parallel","snow","parallelDist","affy","GEOquery","ggplot2","limma") #,"devtools")
if (!require("EWCE"))  devtools::install_github("nathanskene/EWCE")
library(EWCE)

# First, read in the mapping of cell names to cell types
no_cores <- detectCores()
cl <- makeCluster(no_cores)
file="/Users/natske/Box Sync/Single Cell Datasets/Sten Whole Brain 10xGenomics 2018/l5_all.loom"
h5f = H5Fopen(file)
annotDF = data.frame(Cellid=h5f$col_attrs$CellID,level1=h5f$col_attrs$TaxonomyRank4,level2=h5f$col_attrs$ClusterName,desc=h5f$col_attrs$Description)

# Load in the ventral midbrain LOOM file and extract the data.
file="/Users/natske/Box Sync/Single Cell Datasets/Sten Whole Brain 10xGenomics 2018/l1_midbrainventral.loom"
h5f = H5Fopen(file)
exp = t(h5f$matrix)
presentCells = h5f$col_attrs$CellID
colnames(exp) = presentCells
annotDF2 = annotDF[annotDF$Cellid %in% presentCells,]
numCells = table(annotDF[annotDF$Cellid %in% presentCells,]$desc)
annotDF3 = annotDF2[annotDF2$desc %in% names(numCells[numCells>10]),]
dropCellTypes = c("Afferent nuclei of cranial nerves III-V","Cholinergic neurons, midbrain red nucleus","Dopaminergic neurons, periaqueductal grey","Glutamatergic projection neurons of the raphe nucleus","Excitatory neurons, thalamus","Serotonergic neurons, hindbrain")
annotDF4 = annotDF3[!annotDF3$desc %in% dropCellTypes,]
rownames(annotDF4) = annotDF4$Cellid

# Match exp and annot
exp2 = exp[,annotDF4$Cellid]
rownames(exp2) = h5f$row_attrs$Gene
annotLevels = list(level1class=annotDF4$desc,level2class=annotDF4$level2)

# Drop uninformative genes
exp_DROPPED = drop.uninformative.genes(exp=exp2,level2annot = annotLevels$level2class)

# Generate cell type data
fNames_VentralMidbrain10x = generate.celltype.data(exp=exp_DROPPED,annotLevels=annotLevels,groupName="VentralMidbrain10x")
```

Don't actually bother running the above, just use:

```{r }
#load("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017/Data/CellTypeData_VentralMidbrain10x.rda")
data(ctd)
```


# Moran dataset

The Moran dataset is available through GEO using accession number GSE8397. The CEL files should be downloaded directly through the website (GSE8397_RAW.tar). It contains data on superior frontal gyrus and also two regions of the Substantia Nigra: medial and lateral. The data is split across two arrays for each sample (U133A and U133B arrays): these should be normalised seperately and then joined before differential expression analysis.

Moran (GSE8397) used Affymetrix HG_U133 (15 samples of medial parkinsonian SN, 9 samples of lateral parkinsonian SN, 8 medial nigra control samples and 7 lateral nigra control samples). This data is associated with: Moran LB, Duke DC, Deprez M, Dexter DT et al. Whole genome expression profiling of the medial and lateral substantia nigra in Parkinson's disease. Neurogenetics 2006 Mar;7(1):1-11. PMID: 16344956

## Load the data

```{r}
#source("http://bioconductor.org/biocLite.R")
#biocLite("GEOquery")
gse <- getGEO("GSE8397", GSEMatrix = TRUE)
show(gse)
title = as.character(pData(gse[[1]])$title)
dx = rep("Control",length(title)); dx[grep("disease case",title)] = "Parkinsons"
region = rep("Superior Frontal Gyrus",length(title)); region[grep("Lateral",title)] = "Lateral SNc";  region[grep("Medial",title)] = "Medial SNc"
newTitle = sprintf("%s %s",region,dx)
ageGender = pData(gse[[1]])$`age:ch1`
age = as.numeric(gsub("\\;.*","",ageGender))
gender = rep("M",length(age)); gender[grepl("F",ageGender)] = "F"
gsm_id = gsub(".*nnn/|/suppl.*","",as.character(pData(gse[[1]])$supplementary_file))

# Which of the files are associated with U133A and U133B?
u133a_files = gsub(".*/","",as.character(pData(phenoData(gse[[1]]))$supplementary_file)) # U133A files
u133b_files = gsub(".*/","",as.character(pData(phenoData(gse[[2]]))$supplementary_file)) # U133B files

### LOAD THE EXPRESSION DATA ######
#biocLite("affy")
data_u133a <- ReadAffy(filenames=u133a_files,celfile.path="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/GSE8397_Raw",compress=TRUE) # Read in the CEL files then normalize
eset_u133a <- rma(data_u133a)
all_dat_u133a = exprs(eset_u133a)
data_u133b <- ReadAffy(filenames=u133b_files,celfile.path="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/GSE8397_Raw",compress=TRUE) # Read in the CEL files then normalize
eset_u133b <- rma(data_u133b)
all_dat_u133b = exprs(eset_u133b)
all_dat = rbind(all_dat_u133a,all_dat_u133b)
colnames(all_dat) = newTitle
```

## Get the probe annotations from biomaRt

```{r }
library("biomaRt")
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
attrib_hum = listAttributes(human)
annotA = getBM(attributes=c("affy_hg_u133a","external_gene_name"), mart=human)
colnames(annotA)[1] = "probeID"
annotB = getBM(attributes=c("affy_hg_u133b","external_gene_name"), mart=human)
colnames(annotB)[1] = "probeID"
annot  = unique(rbind(annotA,annotB))
annot  = annot[annot$probeID!="",]
dup_probes = unique(annot$probeID[duplicated(annot$probeID)])
annot2 = annot[!(annot$probeID %in% dup_probes),]
colnames(annot2)[2] = "HGNC_Symbol"
```

Now perform the differential expression analysis for each region seperately

```{r }

getTT <- function(coef,eb,probe_annot){
    tt_dx = topTable(eb, coef=coef, adjust="BH",number=1000000)
    colnames(tt_dx)[1] = "probeID"
    tt_dx2 = merge(tt_dx,probe_annot,by="probeID")
    tt_dx2 = tt_dx2[order(tt_dx2$P.Value),]
    return(tt_dx2)
}
plotEWCE <- function(joint_results,labelOrders,fName,width=8,height=10){
    labelOrders = labelOrders[labelOrders %in% joint_results$CellType]
    joint_results$CellType = factor(as.character(joint_results$CellType),levels = labelOrders)
    pdf(file=fName,width=width,height=height)
    print(ewce.plot(joint_results))
    dev.off()        
}

for(reg in unique(region)){
    annot = data.frame(age=age[region==reg],gender=gender[region==reg],dx=dx[region==reg])
    exp   = all_dat[,region==reg]
    
    mod  = model.matrix(~annot$age+annot$gender+annot$dx)
    fit = lmFit(exp,mod)	
    eb = eBayes(fit)
    tt_PD = getTT(coef="annot$dxParkinsons",eb=eb,probe_annot=annot2)
    write.csv(tt_PD,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/TopTables/Moran/dx_%s.csv",reg))
    tt_Gender = getTT(coef="annot$genderM",eb=eb,probe_annot=annot2)
    write.csv(tt_Gender,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/TopTables/Moran/genderMale_%s.csv",reg))
    
    ## Run EWCE
    colnames(tt_PD)[colnames(tt_PD)=="HGNC_Symbol"] = "HGNC.symbol"
    #tt_results = ewce_expression_data(sct_data=ctd,tt=tt_PD,annotLevel=2,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
    #dendroorder = read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderSten.txt",header = FALSE,stringsAsFactors = FALSE)[,1]
    #plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/dx_%s_lvl2.pdf",reg))
    
    tt_results = ewce_expression_data(sct_data=ctd,tt=tt_PD,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
    dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
    plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/dx_%s_lvl1.pdf",reg))
    res_moran = tt_results$joint_results
    res_moran$list = sprintf("Moran et al \\(2006\\)\n%s",reg)
    save(res_moran,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/Moran_%s.rda",reg))
    
    # Run analysis for gender
    colnames(tt_Gender)[colnames(tt_Gender)=="HGNC_Symbol"] = "HGNC.symbol"
    #tt_results = ewce_expression_data(sct_data=ctd,tt=tt_Gender,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
    #dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
    #plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/gender_%s_lvl1.pdf",reg))
}
```

# Lesnick dataset

Lesnick data is available from GEO with accession GSE7621.

Substantia nigra tissue from postmortem brain of normal and Parkinson disease patients were used for RNA extraction and hybridization on Affymetrix U133 Plus 2.0 microarrays: 9 replicates for the controls and 16 replicates for the Parkinson's disease patients were used. Both cohorts included males and females.

```{r}
gse <- getGEO("GSE7621", GSEMatrix = TRUE)
show(gse)
title = as.character(pData(gse[[1]])$title)
dx = rep("Control",length(title)); dx[grep("PD",title)] = "Parkinsons"
gender = as.character(pData(gse[[1]])$`characteristics_ch1.1`)
filenames = gsub(".*/","",as.character(pData(phenoData(gse[[1]]))$supplementary_file))
lesnick_annot = data.frame(fname=filenames,dx=dx,gender=gender)
rownames(lesnick_annot) = annot$fname

### LOAD THE EXPRESSION DATA ######
data <- ReadAffy(celfile.path="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/GSE7621_RAW",compress=TRUE) # Read in the CEL files then normalize
eset <- rma(data)
all_data = exprs(eset)
#colnames(all_dat) = newTitle
#colnames(all_dat) = title
```

## Get the probe annotations from biomaRt

```{r }
library("biomaRt")
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
attrib_hum = listAttributes(human)
annot = getBM(attributes=c("affy_hg_u133_plus_2","external_gene_name"), mart=human)
colnames(annot)[1] = "probeID"
annot  = annot[annot$probeID!="",]
dup_probes = unique(annot$probeID[duplicated(annot$probeID)])
annot2 = annot[!(annot$probeID %in% dup_probes),]
colnames(annot2)[2] = "HGNC_Symbol"
```

Now run the linear model and EWCE analysis

```{r }
getTT_lesnick <- function(coef,eb,probe_annot){
    tt_dx = topTable(eb, coef=coef, adjust="BH",number=1000000)
    tt_dx$probeID=rownames(tt_dx)
    tt_dx2 = merge(tt_dx,probe_annot,by="probeID")
    tt_dx2 = tt_dx2[order(tt_dx2$P.Value),]
    return(tt_dx2)
}

mod  = model.matrix(~lesnick_annot$gender+lesnick_annot$dx)
fit = lmFit(all_data,mod)	
eb = eBayes(fit)
tt_PD = getTT_lesnick(coef="lesnick_annot$dxParkinsons",eb=eb,probe_annot=annot2)
write.csv(tt_PD,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/TopTables/Lesnick/dx.csv"))
tt_Gender = getTT_lesnick(coef="lesnick_annot$gendermale",eb=eb,probe_annot=annot2)
write.csv(tt_Gender,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/TopTables/Lesnick/genderMale.csv"))

## Run EWCE
colnames(tt_PD)[colnames(tt_PD)=="HGNC_Symbol"] = "HGNC.symbol"
tt_results = ewce_expression_data(sct_data=ctd,tt=tt_PD,annotLevel=2,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
dendroorder = read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderSten.txt",header = FALSE,stringsAsFactors = FALSE)[,1]
plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Lesnick/dx_%s_lvl2.pdf",reg),width=15)

tt_results_lesnick = ewce_expression_data(sct_data=ctd,tt=tt_PD,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
res_lesnick = tt_results_lesnick$joint_results
res_lesnick$list = "Lesnick et al (2007)"
save(res_lesnick,file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Lesnick/Lesnick.rda")
dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
plotEWCE(joint_results=tt_results_lesnick$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Lesnick/dx_%s_lvl1.pdf",reg))

# Run analysis for gender
colnames(tt_Gender)[colnames(tt_Gender)=="HGNC_Symbol"] = "HGNC.symbol"
tt_results = ewce_expression_data(sct_data=ctd,tt=tt_Gender,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Lesnick/gender_%s_lvl1.pdf",reg))
```

# VandeBerg dataset

This is the Braak staged Parkinsons data associated with the 2015 paper, "Evidence for Immune Response, Axonal Dysfunction and Reduced Endocytosis in the Substantia Nigra in Early Stage Parkinson's Disease". It is available from GEO with accession GSE49036.

```{r }
###### LOAD THE BRAAK STAGED PARKINSONS DATA
# https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE49036
# Subjects with incidental Lewy body disease (iLBD) may represent the premotor stage of Parkinson’s disease (PD). To identify molecular 
# mechanisms underlying neuronal dysfunction and alpha--synuclein pathology in the premotor phase of PD, we investigated the transcriptome 
# of post-mortem substantia nigra (SN) of iLBD, PD donors and age-matched controls with Braak alpha--synuclein stage ranging from 0-6. In Braak 
# alpha--synuclein stages 1 and 2, we observed deregulation of pathways linked to axonal degeneration, unfolded protein response (UPR), immune 
# response and endocytosis, including axonal guidance signaling, protein kinase A signaling, mTOR signaling, EIF2 signaling and clathrin-mediated 
# endocytosis. In Braak stages 3 and 4, we observed a deregulation in pathways involved in protein translation and cell survival, including mTOR and 
# EIF2 signaling. In Braak stages 5 and 6, we observed deregulation of pathways such as dopaminergic signaling, axonal guidance signaling and thrombin 
# signaling. Throughout the progression of PD pathology, we observed a deregulation of mTOR, EIF2 and regulation of eIF4 and p70S6K signaling in the SN.
# This implicates that molecular mechanisms related to UPR, axonal dysfunction, endocytosis and immune response are an early event in PD pathology, 
# and may hold the key to altering the disease progression in PD.

data <- ReadAffy(celfile.path="/Users/natske/OneDrive - Imperial College London/Datasets/Disease Transcriptome Datasets/Parkinsons/GSE49036_RAW") # Read in the CEL files
eset <- rma(data)
#write.exprs(eset,file="data.txt") # Save data to file (Data is log2 transformed and normalized)
#exprs = exprs(eset)
all_dat = exprs(eset)
colnames(all_dat) = gsub(".CEL","",colnames(all_dat))


##################################
### GET THE PROBE ANNOTATIONS ####

library("biomaRt")
human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
attrib_hum = listAttributes(human)
annot = getBM(attributes=c("affy_hg_u133a_2","external_gene_name"), mart=human)
dup_probes = annot$affy_hg_u133a_2[duplicated(annot$affy_hg_u133a_2)]
annot2 = annot[!(annot$affy_hg_u133a_2 %in% dup_probes),]
annot3 = annot2
colnames(annot3)[2] = "HGNC.symbol"

############################
### LOAD THE ANNOTATION DATA
gse <- getGEO("GSE49036", GSEMatrix = TRUE)
show(gse)
title = as.character(pData(gse[[1]])$title)
dx = rep("Parkinsons",length(title)); dx[grep("Stage 0",title)] = "Control"
rin = as.numeric(pData(gse[[1]])$`rin:ch1`)
braak = as.character(pData(gse[[1]])$`braak stage:ch1`)
gender = all_dat["224588_at",]<8 # There is a clear seperation between genders based on Xist... gender is not otherwise included but this is a standard gender biomarker
filenames = gsub(".*/","",as.character(pData(phenoData(gse[[1]]))$supplementary_file))
annot = data.frame(fname=filenames,dx=dx,gender=gender)
rownames(annot) = annot$fname

## Plot with tSNE to see if gender is a confound
library(Rtsne) # Load package
iris_unique <- unique(iris) # Remove duplicates
set.seed(42) # Sets seed for reproducibility
tsne_out <- Rtsne(t(as.matrix(all_dat)),num_threads=7,perplexity=4) # Run TSNE
tsneData = data.frame(tsne1=tsne_out$Y[,1],tsne2=tsne_out$Y[,2],braak=braak,rin=rin,xist=all_dat["224588_at",])
ggplot(tsneData)+geom_point(aes(x=tsne1,y=tsne2,col=xist))

#sampleAnnot = read.csv("GSE49036_series_matrix.txt")
sAnn = data.frame(chip_id=colnames(all_dat),dx=dx,gender=gender,rin=rin)
sAnn$br_12 = sAnn$br_34 = sAnn$br_56 = rep(0,dim(sAnn)[1])
sAnn$br_12[grep("BR12",title)]=1
sAnn$br_34[grep("BR34",title)]=1
sAnn$br_56[grep("BR56",title)]=1

# Try surrogate variable analysis
library(sva)
mod  = model.matrix(~sAnn$br_12+sAnn$br_34+sAnn$br_56+sAnn$gender+sAnn$rin)
n.sv = num.sv(all_dat,mod,method="leek") # Says there are no surrogate variables
 
# Now do the linear model
fit = lmFit(all_dat,mod)	
eb = eBayes(fit)
tt_12 = topTable(eb, coef="sAnn$br_12", adjust="BH",number=1000000)
tt_34 = topTable(eb, coef="sAnn$br_34", adjust="BH",number=1000000)
tt_56 = topTable(eb, coef="sAnn$br_56", adjust="BH",number=1000000)
tt_12$affy_hg_u133a_2 = rownames(tt_12)
tt_34$affy_hg_u133a_2 = rownames(tt_34)
tt_56$affy_hg_u133a_2 = rownames(tt_56)
tt_12 = merge(tt_12,annot3,by="affy_hg_u133a_2"); tt_12=tt_12[order(tt_12$P.Value),]
tt_34 = merge(tt_34,annot3,by="affy_hg_u133a_2"); tt_34=tt_34[order(tt_34$P.Value),]
tt_56 = merge(tt_56,annot3,by="affy_hg_u133a_2"); tt_56=tt_56[order(tt_56$P.Value),]

tt_results_12 = ewce_expression_data(sct_data=ctd,tt=tt_12,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017/Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
plotEWCE(joint_results=tt_results_12$joint_results,labelOrders=dendroorder,fName="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/VandeBerg/dx_Braak12_lvl1.pdf",height=7)
tt_results_12$joint_results$list = "Braak stage 1-2"

tt_results_34 = ewce_expression_data(sct_data=ctd,tt=tt_34,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
plotEWCE(joint_results=tt_results_34$joint_results,labelOrders=dendroorder,fName="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/VandeBerg/dx_Braak34_lvl1.pdf",height=7)
tt_results_34$joint_results$list = "Braak stage 3-4"

tt_results_56 = ewce_expression_data(sct_data=ctd,tt=tt_56,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
plotEWCE(joint_results=tt_results_56$joint_results,labelOrders=dendroorder,fName="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/VandeBerg/dx_Braak56_lvl1.pdf",height=7)
tt_results_56$joint_results$list = "Braak stage 5-6"

write.csv(tt_12,file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017/Results/TopTables/VandeBerg/tt_12.csv")
write.csv(tt_34,file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017/Results/TopTables/VandeBerg/tt_34.csv")
write.csv(tt_56,file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017/Results/TopTables/VandeBerg/tt_56.csv")

res_merged = rbind(tt_results_12$joint_results,tt_results_34$joint_results,tt_results_56$joint_results)
save(res_merged,file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/VandeBerg/VandeBerg_merged.rda")
plotEWCE(joint_results=res_merged,labelOrders=dendroorder,fName="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/VandeBerg/dx_merged_lvl1.pdf",height=7)
```

# Corradini dataset

GSE43490

```{r}
gse <- getGEO("GSE43490", GSEMatrix = TRUE)
show(gse)
title = as.character(pData(gse[[1]])$title)
dx = pData(gse[[1]])$`disease state:ch1`
dx = gsub("Parkinson's disease","Parkinsons",dx)
region = gsub("brain, ","",as.character(pData(gse[[1]])$`source_name_ch1`))
newTitle = sprintf("%s %s",region,dx)
age = as.numeric(pData(gse[[1]])$`age:ch1`)
gender = as.character(pData(gse[[1]])$`gender:ch1`)
gsm_id = gsub(".*nnn/|/suppl.*","",as.character(pData(gse[[1]])$supplementary_file))

# Get processed expression data directly from the GEO call
exprs = exprs(gse[[1]])
```

Get the probe annotations from GEO. This file was downloaded by clicking on the array reference on GEO.

```{r }
probeAnnot = read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/GPL6480-9577.txt",stringsAsFactors = FALSE,sep="\t",skip=17)[,c("ID","GENE_SYMBOL")]
colnames(probeAnnot)[1] = "probeID"
colnames(probeAnnot)[2] = "HGNC_Symbol"
```

Run differential expression

```{r }
for(reg in unique(region)){
    corradini_annot = data.frame(age=age[region==reg],gender=gender[region==reg],dx=dx[region==reg])
    exp_corradini   = exprs[,region==reg]
    
    mod  = model.matrix(~corradini_annot$age+corradini_annot$gender+corradini_annot$dx)
    fit = lmFit(exp_corradini,mod)	
    eb = eBayes(fit)
    tt_PD = getTT_lesnick(coef="corradini_annot$dxParkinsons",eb=eb,probe_annot=probeAnnot)
    write.csv(tt_PD,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/TopTables/Corradini/dx_%s.csv",reg))
    tt_Gender = getTT_lesnick(coef="corradini_annot$genderMale",eb=eb,probe_annot=probeAnnot)
    write.csv(tt_Gender,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/TopTables/Corradini/genderMale_%s.csv",reg))
    
    ## Run EWCE
    colnames(tt_PD)[colnames(tt_PD)=="HGNC_Symbol"] = "HGNC.symbol"
    #tt_results = ewce_expression_data(sct_data=ctd,tt=tt_PD,annotLevel=2,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
    #dendroorder = read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderSten.txt",header = FALSE,stringsAsFactors = FALSE)[,1]
    #plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/dx_%s_lvl2.pdf",reg))
    
    tt_results = ewce_expression_data(sct_data=ctd,tt=tt_PD,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
    dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
    plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Corradini/dx_%s_lvl1.pdf",reg))
    res_Corradini = tt_results$joint_results
    res_Corradini$list = sprintf("Corradini et al (2014)\n%s",reg)
    save(res_Corradini,file=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Corradini/Corradini_%s.rda",reg))
    
    # Run analysis for gender
    colnames(tt_Gender)[colnames(tt_Gender)=="HGNC_Symbol"] = "HGNC.symbol"
    #tt_results = ewce_expression_data(sct_data=ctd,tt=tt_Gender,annotLevel=1,ttSpecies="human",sctSpecies="mouse",reps=10000,thresh=500)
    #dendroorder = unique(read.csv("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Data/DendroOrderStenDesc.txt",header = FALSE,stringsAsFactors = FALSE,sep="!")[,1])
    #plotEWCE(joint_results=tt_results$joint_results,labelOrders=dendroorder,fName=sprintf("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/gender_%s_lvl1.pdf",reg))
}
```

# Plot all EWCE results together

```{r }
load(file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/VandeBerg/VandeBerg_merged.rda") # res_merged
load(file="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Lesnick/Lesnick.rda") # res_lesnick
load("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/Moran_Medial SNc.rda") # res_moran
res_moran_medial = res_moran
load("/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/Moran/Moran_Lateral SNc.rda")
res_moran_lateral = res_moran
allRes = rbind(res_merged,res_lesnick,res_moran_medial,res_moran_lateral)
plotEWCE(joint_results=allRes,labelOrders=dendroorder,fName="/Users/natske/OneDrive - Imperial College London/Parkinsons_SGN2017//Results/EWCE/dx_merged_lvl1.pdf",height=7)
```
